---
title: ""
output: html_document
---

Peer Assignment Number 1
========================

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up.
These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks.


But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.


This assignment makes use of data from a personal activity monitoring device.
This device collects data at 5 minute intervals through out the day.
The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.The data for this can be downloaded from the web site: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)


The variables included in this dataset are:

1. steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)

2. date: The date on which the measurement was taken in YYYY-MM-DD format

3. interval: Identifier for the 5-minute interval in which measurement was taken

We load the data by using:
```{r}
setwd('D:/')
df<-read.csv("activity.csv")
head(df)
```

The variable 'date' is of type 'factor'. To convert it to type 'Date', I have used the following code.Also i wanted to convert the variable 'interval' into 'factor' format
```{r}
df[,'date']=as.Date(df[,'date'])
df[,3]<-as.factor(df[,3])
class(df[,'date'])
class(df[,'interval'])
```

Now to get an idea about the total number of steps taken each day we plot a histogram
```{r}
hist(tapply(df[,'steps'],df[,'date'],sum),breaks=61,main='Histogram of number of Steps in each day',xlab="No. of Steps")
```

To calculate the mean of number of steps taken each day,we use
```{r}
tapply(df[,'steps'],df[,'date'],mean)
```

To calculate the median of number of steps taken each day,we use
```{r}
tapply(df[,'steps'],df[,'date'],median)
```

To calculate The Mean of Steps taken I used a plot. In the first statement i wanted to remove all the rows with missing values.
```{r}
df1<-df[complete.cases(df),]
plot(levels(df1[,3]),tapply(df1[,1],df1[,3],mean),type='l',xlab='5 Minute Interval',ylab='Mean of steps taken',main="Daily activity pattern")
```

To calculate The Maximum Mean of Steps taken 
```{r}
s<-which.max(tapply(df1[,1],df1[,3],mean))
tapply(df1[,1],df1[,3],mean)[s]
```

To calculate Number of Missing values
```{r}
sum(!complete.cases(df))
```

We replace the missing value by taking mean of values from the 'interval' variable they belong to
```{r}
library(zoo)
df2<-df
df2[,1]<-na.aggregate(df2[,1],df2[,3])
head(df2)
```


Now to get an idea about the total number of steps taken each day we plot a histogram
```{r}
hist(tapply(df2[,'steps'],df2[,'date'],sum),breaks=61,main='Histogram of number of Steps in each day',xlab="No. of Steps")
```

To calculate the mean of the steps taken each day
```{r}
tapply(df2[,'steps'],df2[,'date'],mean)
```

To calculate the median of the steps taken each day
```{r}
tapply(df2[,'steps'],df2[,'date'],median)
```

To make weekends factor
```{r}
p<-function(x,l=c())
{
        s=c('Saturday','Sunday')
        if(x %in% s)
        {
            x='Weekend'    
        }
        else
            x='weekday'
        x
}

mop<-weekdays(df[,2])
d<-lapply(mop,p)
vec<-vector()
df2['Days']<-sapply(d,append,vec)
df2[,'Days']<-as.factor(df2[,'Days'])
head(df2)
```

A panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
library(data.table)
library(lattice)
DT <- data.table(df2)
DS<-DT[,mean(steps),by=list(interval,Days)]
at <- seq(1, nrow(DS), length.out=10)
labels <- round(seq(5, 2355, length.out=10))
xyplot(V1~interval|Days,data=DS,layout=c(1,2),type='l',xlab='Time interval',ylab='Number of steps',main='5-minute interval and the average number of steps taken',scales=list(x=list(at=at, labels=labels, rot=90)))
```
